Viscous Erosion in Porous Media

### 2018-04-04 docs/methods/erosion_methods.tex
NM

In Fig.~\ref{fig:06bodies}, we also show the vorticity field of the surrounding flow. Not only does this field help visualize regions of high shear, but it also allows us to identify certain preferred channels in which the flow is most intense.
As seen in the figure, the flow appears to be concentrated primarily within a single channel that runs horizontally below bodies 1 \& 2 and above bodies 3, 5, \& 6. This main channel widens the fastest, and secondary channels (e.g.~between bodies 1 \& 2 or 3 \& 5 or 5 \& 6) follow. Notice that, in these simulations, the overall vorticity magnitude increases in time due to $\umax$ increasing from keeping the pressure drop fixed.

@@

To further quantify these observations, Fig.~\ref{fig:mbodyplots} shows several measurements taken from the multibody simulations. In all cases, the black curves correspond to the 6-body simulation and gray curves correspond to the 50-body simulation that was shown in Fig.~\ref{fig:50bodies}. First, Fig.~\ref{fig:mbodyplots}(a) shows the area fraction of the solid bodies versus time (i.e.~the 2D counterpart of the volume fraction).

@@

We further quantify many of these observations in Fig.~\ref{fig:mbodyplots}. First, Fig.~\ref{fig:mbodyplots}(a) shows area fraction versus time. The area fraction is the total area of all bodies as compared to the domain area, i.e.~the 2D counterpart of volume fraction. The initial configuration of 6 circles comprises about 40\% of the domain area, and this percentage steadily decreases as the bodies erode. Figure \ref{fig:mbodyplots}(b) shows the inlet/outlet flow-rate, $\umax$, over time. Since we have fixed the pressure drop across the cell in these simulations, $\umax$ steadily increases as the bodies diminish and provide less resistance. Here, the vertical axis is logarithmic, and so the plot suggests exponential increase of $\umax$ at early times. This rapid increase in flow is likely due to the erosion-induced channelization observed above. In Fig.~\ref{fig:mbodyplots}(c) we show the total drag, summed over all bodies, versus time. Though the bodies diminish due to erosion, the steady increase of $\umax$ causes the total drag to remain relatively constant. Only near the end of the simulation does the drag decrease rapidly to zero. To remove the effect of increasing flow rate, we show in Fig.~\ref{fig:mbodyplots}(d) the total drag scaled on $\umax$. This value is a measure for the aggregate {\em resistance} of all bodies. Due to linearity of the Stokes equations, this measure does not depend on the flow-rate $\umax$, only on the bodies and their configuration. As seen here, the resistance decreases rapidly as the bodies erode. Since the vertical axis is logarithmic, the plot suggests an exponential decrease in resistance at early times, which is likely connected to the channelization process.

>>>>

In Fig.~\ref{fig:06bodies}, we also show the vorticity field of the surrounding flow. This field helps not only to visualize regions of high shear, but also to identify certain preferred channels in which the flow is most intense.
As seen in the figure, the flow appears to be concentrated primarily within a single channel that runs horizontally below bodies 1 \& 2 and above bodies 3, 5, \& 6. This main channel widens the fastest, and secondary channels (e.g.~between bodies 1 \& 2 or 3 \& 5 or 5 \& 6) follow. Notice that the overall vorticity magnitude increases in time as the flow rate $\umax$ increases due to keeping the pressure drop fixed.

@@

To further quantify our observations, Fig.~\ref{fig:mbodyplots} shows several measurements taken from the multibody simulations. In all cases, the black curves correspond to the 6-body simulation and gray curves to the 50-body simulation from Fig.~\ref{fig:50bodies}. First, Fig.~\ref{fig:mbodyplots}(a) shows the fraction of area occupied by the solid bodies as it varies in time. In the 6-body simulation, the initial configuration comprises about 40\% of the domain area, and this percentage steadily decreases as the bodies erode. The 50-body simulation shows a similar trend. Figure \ref{fig:mbodyplots}(b) shows how the flow rate, $\umax$, changes in time as a result of fixing the pressure drop across the cell. Notice that, in this figure, the vertical axis is logarithmic. Both simulations show the flow rate to increase rapidly, even during the early stages of erosion when the area fraction has not changed that significantly. In fact, due to the logarithmic axis, the nearly linear trend seen in the figures corresponds to exponential growth of $\umax$. The change in area-fraction alone is not enough to account for this rapid growth, implying that it is the way in which erosion reshapes the bodies that allows for high flow rates.

@@

Figure \ref{fig:mbodyplots}(b) shows the inlet/outlet flow-rate, $\umax$, over time. Since we have fixed the pressure drop across the cell in these simulations, $\umax$ steadily increases as the bodies diminish and provide less resistance. Here, the vertical axis is logarithmic, and so the plot suggests exponential increase of $\umax$ at early times. This rapid increase in flow is likely due to the erosion-induced channelization observed above. In Fig.~\ref{fig:mbodyplots}(c) we show the total drag, summed over all bodies, versus time. Though the bodies diminish due to erosion, the steady increase of $\umax$ causes the total drag to remain relatively constant. Only near the end of the simulation does the drag decrease rapidly to zero. To remove the effect of increasing flow rate, we show in Fig.~\ref{fig:mbodyplots}(d) the total drag scaled on $\umax$. This value is a measure for the aggregate {\em resistance} of all bodies. Due to linearity of the Stokes equations, this measure does not depend on the flow-rate $\umax$, only on the bodies and their configuration. As seen here, the resistance decreases rapidly as the bodies erode. Since the vertical axis is logarithmic, the plot suggests an exponential decrease in resistance at early times, which is likely connected to the channelization process.
